---
title: "Summarize"
author: "Avantika_Negi"
date: "2023-01-07"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```

#delay of 12 mins, reduced evrything to one value
```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

#freely omit NA values, departure column has NAs in it. We have to consider exactly what we want in this case
```{r}
summarize(flights, delay = mean(dep_delay))
```

#Going to group by the day of the year, we want to have the mean depaarture delay now.
```{r}
by_day <- group_by(flights, year, month, day)
summarize(by_day, delay = mean(dep_delay, na.rm = TRUE))
```


#microplate reader readings for three technical replicates 
#specify each condition in the three columns, rows and group them by the exerimental condition


```{r}
summarize(flights, delay = n())  #mean(dep_delay, na.rm = TRUE))
```



# Find the realtionship between distance and average delay fro each destination

#group flights by destination

#summarize to compute distance, average delay and number of flgihts

# filter, to remove noisy points and Honululu airport

#Pipe operator, code reads easily
```{r}
delays <- flights %>%
  group_by(dest) %>%
  summarize(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>%
  filter(count > 20, dest != "HNL")
delays
``` 
#The grouping is by destination, not by yr, month. The summarize only gives the columns included in the three lines and the grouping variable (lose a lot of columns). 


```{r}
delays <- flights %>%
  group_by(dest) %>%
  summarize(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE),
    day = first(day)
  ) %>%
  filter(count > 20, dest != "HNL")
delays
``` 


## When using a dataframe function na.rm in r refers to the logical parameter that tells the function whether or not to remove NA values from the calculation. 
## It literally means NA remove. It is neither a function nor an operation. It is simply a parameter used by several dataframe functions. They include colSums(), rowSums(), colMeans() and rowMeans(). When na.rm is TRUE, the function skips over any NA values. 
## However, when na.rm is FALSE, then it returns NA from the calculation being done on the entire row or column.



#Alternative to PIPE
#filter(summarize(group_by(flights), dest), count = ...), count > 20 ...)

```{r}
#making another table
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))

delays <- not_cancelled %>%
  group_by(tailnum) %>%
  summarize(
    delay = mean(arr_delay)
  )
  
ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)

```


```{r}
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))

#writing a table into a ggplot!!!
p <- not_cancelled %>%
   group_by(tailnum) %>%
   summarize(
     delay = mean(arr_delay),
     n = n()
   ) %>%
   
   ggplot(mapping = aes(y = delay, x = n)) + 
     geom_point(alpha = 1/10)
p

```



```{r}
not_cancelled <- flights %>% filter(!is.na(dep_delay), !is.na(arr_delay))

#writing a table into a ggplot!!!
p <- not_cancelled %>%
   group_by(tailnum) %>%
   summarize(
     delay = mean(arr_delay),
     n = n()
   ) %>%
   filter(n > 25) %>%
   ggplot(mapping = aes(y = delay, x = n)) + 
     geom_point(alpha = 1/10)
p

```


#CHAPTER 7.3, 7.7


#Demo histograms
```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
```
#way of looking at probability densities


```{r}
diamonds %>%
  filter(carat < 3) %>%
  ggplot() +
  geom_freqpoly(mapping = aes(x = carat), binwidth = 0.5)
```


#Nice thing about freqpoly allows one to compare 
```{r}
diamonds %>%
  filter(carat < 3) %>%
  ggplot() +
  geom_freqpoly(mapping = aes(x = carat, color = cut), binwidth = 0.5)
```


```{r}
  ggplot(diamonds) +
  geom_freqpoly(aes(carat, color = cut), binwidth = 0.5)
```



