---
title: "5DataTransform"
author: "Avantika_Negi"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nycflights13)
library(tidyverse)
```


#dplyr
##dplyr is a subset library of tidyverse used for manipulating tables
```{r}
flights
```


#basic ops

filter()      #select by values 
arrange()     #reorder rows
select()      #pick by names 
mutate()      #create new vars
summary()     #colllapse to a single val

 - group_by()   #divides tbales into groups 
 
 
 #Filter 
 
```{r}
filter(flights, month == 1, day == 1)
```


# we want either of the two conditions true for the filtering operation
```{r}
filter(flights, month == 1 | day == 1)
```

# to save the work of the filter operation one needs to assign the output to a variable

```{r}
jan1 <- filter(flights, month == 1 | day == 1)
```


```{r}
(jan1 <- filter(flights, month != 9)) #eveything except Sept
```

#Exclusive OR also exists 

```{r}
(jan1 <- filter(flights, !(month >9 & day > 15)))
```

#to have a range for flights in the summer time
#frequently used operator, can be done with greater, less than and and operator but that is long, so %in% operator works so c(6, 7, 8) is useful
#n. a value that does not exist in the table, toxic constant
```{r}
(jan1 <- filter(flights, month %in% c(6, 7, 8)))
```



#NA values 
```{r}
NA > 5
```

```{r}
NA == 5
```


```{r}
NA == NA   #What!!
```


```{r}
NA == NA   #What!!
x <- NA
y <- NA
x == y
``` 
# Well, we don't know this data, hence this weird output

```{r}
NA == NA   #What!!
x <- NA
y <- NA
x == y
is.na(NA)
``` 



#Arrange

```{r}
arrange(flights, year, month, day)
```

```{r}
arrange(flights, year, desc(month), desc(day))
```


```{r}
arrange(flights, year, desc(dep_delay))
```

#Select columns
```{r}
select(flights, year, month, day, dep_delay)
```


#makes it easier to read, if we don't care about year month or days 
```{r}
select(flights, -year, -month, -day)
```


```{r}
select(flights, -year:day)
```



```{r}
select(flights, -(year:day))
```



#Mutate
```{r}
flights_small <- select(flights,
                        year:day, 
                        ends_with("delay"),
                        distance,
                        air_time)

mutate(flights_small,
       gain = dep_delay - arr_delay,
       speed = distance /air_time * 60)

```


```{r}
flights_small <- select(flights,
                        year:day, 
                        ends_with("delay"),
                        distance,
                        air_time)

mutate(flights_small,
       gain = dep_delay - arr_delay,
       speed = distance /air_time * 60)
       hours = air_time / 60
       gain_per_hour = gain / hours)

```
