---
title: "Lecture_5a_Tidy_Data_5b_factors"
author: "Avantika_Negi"
date: "2023-01-16"
output: html_document
---
## Arranging data appropriately so that it can plug into tiderverse and ggplot and other functions. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```
#Chapter 12 from R4DS
###know the situations when these manipulations on tidying dats are not needed 


## 1. Each variable must have its own column
## 2. Each observation it its own row
## 3. Each value in its own cell. 

```{r}
##Series of tidyverse tables 
table1

# vectorization: define a function on one of the rows and tell it to apply to a column in the same  way since the data type is same.
# Always have copies of your data set 

```



```{r}
table4a

#each column is a year 
## We want each year as a column, reshape this to look like the table above
## Squeeze it to a longer table using pivot 
table4a %>%
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")

```

```{r}
tidy4a <- table4a %>%
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
tidy4a
```



```{r}
tidy4b <- table4a %>%
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
tidy4b
```

```{r}
##joining these tables 4a and 4b togethe: merging country, yr 


tidy4 <- left_join(tidy4a, tidy4b)
ggplot(tidy4, aes(cases, population)) + 
  geom_point()

```


```{r}
table2
## does not really look tidy 
## aesthetics are columns, cant do the previous plot with this table
```






```{r}
table2
ggplot(table2, aes(year, count)) + geom_point()
  
## We need to spread these values out 
```



```{r}
table2 %>%
  pivot_wider(names_from = type, values_from = count)
```
  
  
  
```{r}
## Single cell has more than one value in it
## Split this column into two new populations 
table3
table3 %>%
  separate(rate, into = c("cases", "populations"))
```

# Look at chapter 14 and 16 in R4DS on Strings 
```{r}
## Single cell has more than one value in it
## Split this column into two new populations 
table3

## separators is useful for seraching amongst strings 
table3 %>%
  separate(rate, into = c("cases", "populations"), sep = "foo")
```

```{r}
# nycflights table and separate it to hours and minutes 
nycflights13::flights %>%
  select(year, month, day, dep_time) %>%
  separate(dep_time, into = c("hour", "minute"), sep = 2)
## interpreted as charcter positions to split at 
```


```{r}
table5 %>%
  unite(new, century, year, sep = "")

```

#Chapter 15: FACTORS
### Small tables examples: look at the exercises at the end of Chapter 12 

```{r}
#install.packages('XQuartz', repos = "www.xquartz.org")
x1 <- c("Dec", "Apr", "Jan", "Mar")
x2 <- c("Dec", "Apr", "Jam", "Mar")
sort(x1)
##converting strings to factors

month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
y2 <- parse_factor(x2, levels =  month_levels)
View(gss_cat)
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )
relig_summary

ggplot(relig_summary, aes(tvhours, relig)) +
  geom_point()
##But we cant really interpret anything from  this plot, we need more structure to this. 

##factoring ordering functions are handy here 
###tvhours is a continuous variable a no. not a factor, religion is interesting to play with 
ggplot(relig_summary, aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point()
```
###greatest to least sorted:
###JUST FOR THE Y-AXIS LABEL HERE 
###Renamed the column of the table in the correct order and piped that into ggplot 
```{r}

relig_summary %>%
  mutate(relig = fct_reorder(relig, tvhours)) %>%
  ggplot(relig_summary, mapping = aes(tvhours, relig)) +
  geom_point()


```

```{r}
rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

##Income Summary: ordering on the basis of age  
ggplot(rincome_summary, aes(age, rincome, age)) + 
  geom_point()

ggplot(rincome_summary, aes(age, fct_reorder(rincome, age))) + geom_point()
##Keeping the not applicable lable at the bottom 
ggplot(rincome_summary, aes(age, fct_relevel(rincome, "Not applicable"))) + geom_point()


```



##Getting rid of unknown ages and counting the number of each by marital status, the grouping the ages and calculating the proportion
```{r}
by_age <- gss_cat %>%
  filter(!is.na(age)) %>%
  count(age, marital) %>%
  group_by(age) %>%
  mutate(prop = n / sum(n))
by_age

ggplot(by_age, aes(age, prop, colour = marital)) +
  geom_line(na.rm = TRUE) ##removing NAs 

##cant tell the fiff between not married and never married Kovar Lab
###trend lines: key mathcing up with the end lines: closest data, makes it easier to see the trend without looking at the graph/data 
##two things have to be reordered: need x and y value to order, hence reorder2
ggplot(by_age, aes(age, prop, colour = fct_reorder2(marital, age, prop))) +
  geom_line() +
  labs(colour = "marital")
```



```{r}

##sorting by freq and reverse it 
gss_cat %>%
  mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
  ggplot(aes(marital)) +
    geom_bar()

```


#VERY USEFUL
```{r}
##labels evolve with experiments, or maybe typos, not settled on common conventions
###the coding of the factors in thatb case has to be changed 

gss_cat %>% count(partyid)
#change the column with recoded factors
gss_cat %>%
  mutate(partyid = fct_recode(partyid,
    "Republican, strong"    = "Strong republican",
    "Republican, weak"      = "Not str republican",
    "Independent, near rep" = "Ind,near rep",
    "Independent, near dem" = "Ind,near dem",
    "Democrat, weak"        = "Not str democrat",
    "Democrat, strong"      = "Strong democrat",
    "Other"                 = "No answer",
    "Other"                 = "Don't know",
    "Other"                 = "Other party"
  )) %>%
  count(partyid)
##can combine too, if tring isnt imp, by changing label names 

```


