---
title: "Lecture_6a_Functions"
author: "Avantika_Negi"
date: "2023-01-19"
output: html_document
---
#GO THROUGH Chapter 19 in R4DS 
#Chapter 17, 18-21

# Functions
# When to use?
# * When you have cut-and-pasted three times, it is good to start constructing a function to do the job
# 1. Named, so easier label to understand
# 2. Only one place to edit 
# 3. Cut & paste errors go away

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

##```{ r name-here}

##helpful for constructing models 

##several lines of code

### one function or operation, function; jump back and forth between code blocks a

df <- tibble::tibble(
  a = rnorm(10), #gaussian distribution
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
##rescale the vectors so that the values lie between 0 and 1
##one can name functions in R
rescale01 <- function(x)  {
  xmin <- min(x, na.rm = TRUE)
  xmax <- max(x, na.rm = TRUE)
  (x - xmin) / (xmax - xmin)
}

rescale01(c(-10, 0, 10))

df$a <- rescale01(df$a)
View(df)
# df$b <- (df$b - min(df$b, na.rm = TRUE)) / 
#   (max(df$b, na.rm = TRUE) - min(df$a, na.rm = TRUE))
# df$c <- (df$c - min(df$c, na.rm = TRUE)) / 
#   (max(df$c, na.rm = TRUE) - min(df$c, na.rm = TRUE))
# df$d <- (df$d - min(df$d, na.rm = TRUE)) / 
#   (max(df$d, na.rm = TRUE) - min(df$d, na.rm = TRUE))
# View(df)

```

#TESTS
```{r}
rescale0m1 <- function(x, rm = TRUE, scale_factor = 1)  {
  xmean <- mean(x, na.rm = rm)
  xmin <- min(x, na.rm = TRUE)
  xmax <- max(x, na.rm = TRUE)
  scale_factor * (x - xmin) / (xmax - xmin)
}
##function call
rescale0m1(c(-10, 0, NA, 10), scale_factor = 20)#rm = FALSE)
```

#Good short argument names 
data: x, y, z #vectors
dataframe: df df_foo
weights: w
indices: i, j
length: n
everything else-- gets a better name



```{r}

test_that("Rescale uniform list", {
  expect_equal(rescale01(c(-10, 0, 10)), c(0.0, 0.3, 1.0))
}
)

```

#Lecture 6b Iteration and Flow Control
##looping, if-then statements

##Conditional execution



if (condition) {
    # Code that runs if it is true 
} else {
#Code that runs if the condition is false 
}


if (condition1) {
    # Code that runs if it is true 
} else if (condition 2) {
#Code that runs if the condition is false 
} else {
#Code that runs if the condition is false 
}

```{r}
#Function that computes weighted statistics for the given data 
wt_mean <- function(x, w) {
  if (length(x) != length(w)) {
    stop("`x` and `w` must be the same length", call. = FALSE)
      
  }
  sum(x * w) / sum(w)
}

#w and x are of diff lengths
wt_mean(1:6, 1:3)


```


```{r}

foo <- function(x) {
  if ( length(x) != 1) {
      stop("`x` must have length 1")
  }
  if ( x == 3 || x ==7) {
    # this calculation is simple
    return(-7)
  }
  # But here is a big calc
  # ...
  # ...
  x
}
foo(10)

```


## Having an explicit return is good because in a long code one forgets where they wrote the function header

```{r}
foo <- function(x) {
  if (x != 0) {
    # But here is a big calc
  # ...
  # ...
  } else {
    # this calculation is simple
    
  }
}
foo(10)
```

#For loops

```{r}

output <- vector("double", ncol(df)) #preallocated output, if possible

for (i in seq_along(df)) { #defining the sequence
    output[[i]] <- median(df[[i]])
}
output

```

for(i in seq_along(df)) #index values i
```{r}

xs <- c("A", "B", "C")
for(x in xs) {
    print(x)
}
```



```{r}
for (nm in names(df))  { 
    print(nm)
    print(median(df[[nm]]))
}

```









