---
title: "Lecture_5a_Tidy_Data"
author: "Avantika_Negi"
date: "2023-01-16"
output: html_document
---
## Arranging data appropriately so that it can plug into tiderverse and ggplot and other functions. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
```
#Chapter 12 from R4DS
###know the situations when these manipulations on tidying dats are not needed 


## 1. Each variable must have its own column
## 2. Each observation it its own row
## 3. Each value in its own cell. 

```{r}
##Series of tidyverse tables 
table1

# vectorization: define a function on one of the rows and tell it to apply to a column in the same  way since the data type is same.
# Always have copies of your data set 

```



```{r}
table4a

#each column is a year 
## We want each year as a column, reshape this to look like the table above
## Squeeze it to a longer table using pivot 
table4a %>%
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")

```

```{r}
tidy4a <- table4a %>%
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "cases")
tidy4a
```



```{r}
tidy4b <- table4a %>%
  pivot_longer(c(`1999`, `2000`), names_to = "year", values_to = "population")
tidy4b
```

```{r}
##joining these tables 4a and 4b togethe: merging country, yr 


tidy4 <- left_join(tidy4a, tidy4b)
ggplot(tidy4, aes(cases, population)) + 
  geom_point()

```


```{r}
table2
## does not really look tidy 
## aesthetics are columns, cant do the previous plot with this table
```






```{r}
table2
ggplot(table2, aes(year, count)) + geom_point()
  
## We need to spread these values out 
```



```{r}
table2 %>%
  pivot_wider(names_from = type, values_from = count)
```
  
  
  
```{r}
## Single cell has more than one value in it
## Split this column into two new populations 
table3
table3 %>%
  separate(rate, into = c("cases", "populations"))
```

# Look at chapter 14 and 16 in R4DS on Strings 
```{r}
## Single cell has more than one value in it
## Split this column into two new populations 
table3

## separators is useful for seraching amongst strings 
table3 %>%
  separate(rate, into = c("cases", "populations"), sep = "foo")
```

```{r}
# nycflights table and separate it to hours and minutes 
nycflights13::flights %>%
  select(year, month, day, dep_time) %>%
  separate(dep_time, into = c("hour", "minute"), sep = 2)
## interpreted as charcter positions to split at 
```


```{r}
table5 %>%
  unite(new, century, year, sep = "")

```




### Small tables examples: look at the exercises at the end of Chapter 12 



